---
title: "Effects of Dietary Restriction and Insulin-Like Growth Factor Supplementation on Lizard Tail Regeneration"
date: "2026-07-26"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.align = "center",
  fig.retina = 2
)
```

# Study Overview

The objective of this study was to determine whether energetic state constrains tail regeneration in adult female brown anoles (*Anolis sagrei*) and whether supplementation with insulin-like growth factor 1 (IGF1) or insulin-like growth factor 2 (IGF2) modifies this relationship.

- Energetic state was quantified as percent body mass loss and analyzed as a continuous predictor of regenerative performance.
- The primary response variable was cumulative percent tail regeneration measured weekly over eight weeks following tail autotomy.
- Secondary analyses evaluated absolute regenerated tail length and weekly regeneration rate.
- Linear mixed-effects models included animal identity as a random effect to account for repeated measurements.

The analysis proceeds in the same order as the manuscript:

1. Verify successful randomization and dietary manipulation.
2. Establish the relationship between energetic state and regeneration.
3. Test whether IGF supplementation modifies that relationship.
4. Evaluate robustness, alternative response variables, dose effects, and late-stage observations.
5. Generate publication figures and supporting outputs.

# 1. Reproducibility Setup

## 1.1 Load packages

```{r packages}
# Data manipulation
library(dplyr)
library(tidyr)

# Statistical modeling
library(lme4)
library(nlme)
library(car)
library(lmtest)
library(MuMIn)
library(bestNormalize)
library(multcomp)
library(emmeans)
library(performance)

# Model predictions and summaries
library(ggeffects)
library(Rmisc)

# Visualization
library(ggplot2)
library(patchwork)
library(plotly)
library(viridis)
library(hrbrthemes)

# Reproducible random-number generation for bootstrap analyses
set.seed(123)
```

## 1.2 Shared plotting objects and color palettes

```{r plot-settings}
# Position adjustments used in several figures
pd <- position_dodge(0.02)
dodge <- position_dodge(width = 0.6)

# Treatment colors used throughout the analysis
treatment_colors <- c(
  "AdLib"   = "grey40",
  "DR"      = "grey70",
  "DR.IGF1" = "#E87722",
  "DR.IGF2" = "#0C2340"
)

igf_colors <- c(
  "No_IGF" = "grey40",
  "IGF1"   = "#E87722",
  "IGF2"   = "#0C2340"
)
```

# 2. Data Import and Curation

## 2.1 Pre-experimental body-mass dataset

```{r import-weight-data}
diet <- read.csv(
  "WL.analysis.csv",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

# The source file contains a trailing space in the Week column name.
names(diet)[trimws(names(diet)) == "Week"] <- "Week"

# Standardize factor order for consistent output.
diet <- diet %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("AdLib", "DR", "DR.IGF1", "DR.IGF2")
    )
  )

str(diet)
```

## 2.2 Eight-week regeneration dataset

```{r import-regeneration-data}
data <- read.csv(
  "R.analysis.currated.csv",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

# Remove observations from four individuals following substantial secondary
# tail loss unrelated to the original experimental autotomy.
#
# The original July analysis removed the complete records for these animals.
# This exclusion is retained here so that the reorganized file reproduces the
# same analytical dataset and results.
excluded_animals <- c("BC121", "BC144", "BC383", "BC561")

data <- data %>%
  filter(!Animal_ID %in% excluded_animals) %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("AdLib", "DR", "DR.IGF1", "DR.IGF2")
    ),
    Diet = case_when(
      Treatment == "AdLib" ~ "AdLib",
      Treatment %in% c("DR", "DR.IGF1", "DR.IGF2") ~ "Restricted",
      TRUE ~ NA_character_
    ),
    Diet = factor(Diet, levels = c("AdLib", "Restricted")),
    IGF = if_else(
      Treatment %in% c("DR.IGF1", "DR.IGF2"),
      "IGF",
      "No_IGF"
    ),
    IGF = factor(IGF, levels = c("No_IGF", "IGF")),
    IGF_type = case_when(
      Treatment %in% c("AdLib", "DR") ~ "No_IGF",
      Treatment == "DR.IGF1" ~ "IGF1",
      Treatment == "DR.IGF2" ~ "IGF2",
      TRUE ~ NA_character_
    ),
    IGF_type = factor(IGF_type, levels = c("No_IGF", "IGF1", "IGF2"))
  )

str(data)
table(data$Treatment, useNA = "ifany")
length(unique(data$Animal_ID))
```

# 3. Verification of Randomization

Animals should not differ systematically in body size or original tail length before treatment implementation.

## 3.1 Baseline subset

```{r baseline-subset}
pre_diet <- diet %>%
  filter(Week == -4)
```

## 3.2 Initial snout–vent length

```{r baseline-svl}
model_svl <- lm(SVL ~ Treatment, data = pre_diet)
anova(model_svl)
summary(model_svl)

plot_svl <- ggplot(
  pre_diet,
  aes(x = Treatment, y = SVL, fill = Treatment)
) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.10, outlier.shape = NA) +
  geom_jitter(width = 0.10) +
  scale_fill_manual(values = treatment_colors) +
  labs(x = "Treatment", y = "Snout–vent length") +
  theme_classic() +
  theme(legend.position = "none")

plot_svl

ggsave(
  filename = "svl.png",
  plot = plot_svl,
  width = 5,
  height = 4,
  dpi = 600
)
```

**Manuscript result:** Initial snout–vent length did not differ among treatment groups, \(F_{3,72}=0.65\), \(p=0.58\).

## 3.3 Initial body mass

```{r baseline-mass}
model_mass <- lm(Mass ~ Treatment, data = pre_diet)
anova(model_mass)
summary(model_mass)

plot_mass <- ggplot(
  pre_diet,
  aes(x = Treatment, y = Mass, fill = Treatment)
) +
  geom_violin(trim = FALSE, scale = "width") +
  geom_boxplot(width = 0.10, outlier.shape = NA) +
  geom_jitter(width = 0.10) +
  scale_fill_manual(values = treatment_colors) +
  labs(x = "Treatment", y = "Initial body mass") +
  theme_classic() +
  theme(legend.position = "none")

plot_mass

ggsave(
  filename = "mass.png",
  plot = plot_mass,
  width = 5,
  height = 4,
  dpi = 600
)
```

**Manuscript result:** Initial body mass did not differ among treatment groups, \(F_{3,72}=0.55\), \(p=0.64\).

## 3.4 Original tail length

```{r baseline-tail}
model_tail_baseline <- lm(Tail ~ Treatment, data = pre_diet)
anova(model_tail_baseline)
summary(model_tail_baseline)
```

**Manuscript result:** Original tail length did not differ significantly among treatment groups, \(F_{3,72}=2.26\), \(p=0.09\).

# 4. Verification of the Dietary Manipulation

## 4.1 Percent body-mass loss after acclimation

```{r diet-manipulation}
diet_final_acclimation <- diet %>%
  filter(Week == -1)

model_diet_loss <- lm(Perc_WL ~ Diet, data = diet_final_acclimation)
anova(model_diet_loss)
summary(model_diet_loss)

diet_group_summary <- diet_final_acclimation %>%
  group_by(Diet) %>%
  summarise(
    n = sum(!is.na(Perc_WL)),
    mean_percent_loss = mean(Perc_WL, na.rm = TRUE),
    sd_percent_loss = sd(Perc_WL, na.rm = TRUE),
    mean_percent_original_mass = mean(Perc_Orig, na.rm = TRUE),
    .groups = "drop"
  )

diet_group_summary
```

**Manuscript result:** Diet-restricted females lost significantly more body mass than ad libitum-fed females, \(F_{1,67}=13.85\), \(p=0.0004\).

## 4.2 Figure 1A: energetic manipulation

```{r figure-diet-manipulation}
plot_diet <- ggplot(
  diet_final_acclimation,
  aes(x = Diet, y = Perc_Orig)
) +
  geom_violin(
    fill = "grey75",
    color = "black",
    trim = FALSE,
    scale = "width",
    alpha = 0.8
  ) +
  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    fill = "white",
    color = "black"
  ) +
  geom_jitter(
    width = 0.08,
    size = 1.8,
    shape = 21,
    fill = "white",
    color = "black",
    alpha = 0.8
  ) +
  labs(
    x = "Diet",
    y = "Percent of original body mass"
  ) +
  theme_classic(base_size = 12)

plot_diet

ggsave(
  filename = "diet.png",
  plot = plot_diet,
  width = 5,
  height = 4,
  dpi = 600
)
```

## 4.3 Body-mass trajectories during acclimation

These plots are retained from the original analysis to document how body mass changed before tail autotomy.

```{r acclimation-trajectories}
plot_mass_by_week_violin <- ggplot(
  diet,
  aes(x = factor(Week), y = Mass, fill = Diet)
) +
  geom_violin(
    trim = FALSE,
    position = dodge,
    scale = "width",
    width = 0.5
  ) +
  geom_boxplot(
    width = 0.10,
    position = dodge,
    outlier.shape = NA
  ) +
  geom_jitter(position = dodge) +
  scale_fill_manual(values = c("Adlib" = "slategrey", "DR" = "lemonchiffon2")) +
  labs(x = "Week", y = "Body mass") +
  theme_classic()

plot_mass_by_week <- ggplot(
  diet,
  aes(x = Week, y = Mass, color = Diet)
) +
  geom_smooth() +
  scale_color_manual(values = c("Adlib" = "slategrey", "DR" = "lemonchiffon2")) +
  theme_classic()

plot_loss_by_week <- ggplot(
  diet,
  aes(x = Week, y = Perc_WL, color = Diet)
) +
  geom_smooth() +
  scale_color_manual(values = c("Adlib" = "slategrey", "DR" = "lemonchiffon2")) +
  theme_classic()

plot_percent_original_by_week <- ggplot(
  diet,
  aes(x = Week, y = Perc_Orig, color = Diet)
) +
  geom_smooth() +
  scale_color_manual(values = c("Adlib" = "slategrey", "DR" = "lemonchiffon2")) +
  theme_classic()

plot_mass_by_week_violin
plot_mass_by_week
plot_loss_by_week
plot_percent_original_by_week

ggsave(
  filename = "diet2.png",
  plot = plot_mass_by_week_violin,
  width = 5,
  height = 4,
  dpi = 600
)

ggsave(
  filename = "diet4.png",
  plot = plot_percent_original_by_week,
  width = 4.75,
  height = 4,
  dpi = 600
)
```

# 5. Establishing the Energetic Trade-off

## 5.1 Primary cumulative-percent-regeneration model

```{r tradeoff-primary-model}
model_weightloss <- lmer(
  Perc.Regeneration ~ Week * Perc.WeightLoss +
    (1 | Animal_ID),
  data = data
)

summary(model_weightloss)
anova(model_weightloss)
```

**Manuscript results:**

- Week: \(\beta=4.86\), \(t=25.31\), \(p<0.0001\)
- Percent body-mass loss: \(\beta=0.074\), \(t=1.39\), \(p=0.165\)
- Week × percent body-mass loss: \(\beta=-0.030\), \(t=-3.01\), \(p=0.0028\)

A negative Week × Percent Weight Loss interaction indicates that animals experiencing greater energetic deficits regenerated progressively less tail tissue over time.

## 5.2 Alternative regeneration metrics

These models retain the same fixed- and random-effect structure while replacing the response variable.

```{r tradeoff-alternative-responses}
model_weightloss_rate <- lmer(
  Rate_Reg ~ Week * Perc.WeightLoss +
    (1 | Animal_ID),
  data = data
)

model_weightloss_absolute <- lmer(
  Regeneration ~ Week * Perc.WeightLoss +
    (1 | Animal_ID),
  data = data
)

summary(model_weightloss_rate)
summary(model_weightloss_absolute)
```

## 5.3 Exploratory visualization by energetic-state category

The primary analyses treat percent body-mass loss continuously. The following three-category plot is retained only as a descriptive visualization.

```{r exploratory-weightloss-categories}
data <- data %>%
  mutate(
    WL_group = cut(Perc.WeightLoss, breaks = 3)
  )

ggplot(
  data,
  aes(
    x = Week,
    y = Perc.Regeneration,
    color = WL_group,
    group = WL_group
  )
) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(fun = mean, geom = "point") +
  labs(
    x = "Week",
    y = "Cumulative percent regeneration",
    color = "Weight-loss category"
  ) +
  theme_classic()
```

# 6. Testing Rescue by IGF Supplementation

## 6.1 Primary IGF1-versus-IGF2 rescue model

```{r rescue-primary-model}
model_igf_split <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * IGF_type +
    Tail +
    (1 | Animal_ID),
  data = data
)

summary(model_igf_split)
anova(model_igf_split)
```

**Manuscript interpretation:** IGF2 significantly altered the increasingly negative relationship between energetic deficit and regeneration, whereas IGF1 did not.

## 6.2 Model-based predictions

```{r rescue-model-predictions}
predicted_all_weeks <- ggpredict(
  model_igf_split,
  terms = c(
    "Perc.WeightLoss",
    "IGF_type",
    "Week [2,4,6,8]"
  )
)

predicted_early_weeks <- ggpredict(
  model_igf_split,
  terms = c(
    "Perc.WeightLoss",
    "IGF_type",
    "Week [2,4]"
  )
)

predicted_all_weeks
predicted_early_weeks
```

```{r rescue-prediction-plot}
plot_predictions <- ggplot(
  predicted_all_weeks,
  aes(
    x = x,
    y = predicted,
    color = group,
    fill = group
  )
) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high),
    alpha = 0.15,
    color = NA
  ) +
  geom_line(linewidth = 1.2) +
  facet_wrap(~facet) +
  scale_color_manual(values = igf_colors) +
  scale_fill_manual(values = igf_colors) +
  labs(
    x = "Percent body-mass loss",
    y = "Predicted cumulative percent regeneration",
    color = "IGF treatment",
    fill = "IGF treatment"
  ) +
  theme_classic()

plot_predictions
```

## 6.3 Week-specific energetic-state slopes

```{r rescue-week-specific-slopes}
week_specific_slopes <- emtrends(
  model_igf_split,
  ~ IGF_type | Week,
  var = "Perc.WeightLoss",
  at = list(Week = c(2, 4, 6, 8))
)

summary(week_specific_slopes, infer = TRUE)
test(week_specific_slopes)

week_specific_slopes_df <- as.data.frame(
  summary(week_specific_slopes, infer = TRUE)
)
```

## 6.4 Pairwise comparisons of slopes

The manuscript describes Tukey-adjusted comparisons among supplementation groups. Both the full pairwise comparisons and treatment-versus-control contrasts are retained.

```{r rescue-pairwise-comparisons}
# All pairwise comparisons within week, Tukey adjusted
pairwise_slopes_tukey <- pairs(
  week_specific_slopes,
  adjust = "tukey"
)

pairwise_slopes_tukey

# Treatment-versus-control contrasts retained from the original figure workflow
contrasts_vs_control <- contrast(
  week_specific_slopes,
  method = "trt.vs.ctrl",
  ref = "No_IGF",
  adjust = "none"
)

contrasts_vs_control
```

## 6.5 Figure 2A: week-specific rescue slopes

```{r figure-rescue-slopes}
comparison_labels <- as.data.frame(contrasts_vs_control) %>%
  mutate(
    IGF_type = case_when(
      grepl("IGF1", contrast) ~ "IGF1",
      grepl("IGF2", contrast) ~ "IGF2",
      TRUE ~ NA_character_
    ),
    sig_vs_control = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.10  ~ "†",
      TRUE            ~ ""
    )
  ) %>%
  select(Week, IGF_type, sig_vs_control)

control_labels <- week_specific_slopes_df %>%
  filter(IGF_type == "No_IGF") %>%
  transmute(
    Week,
    IGF_type,
    sig_vs_control = ""
  )

rescue_plot_data <- week_specific_slopes_df %>%
  left_join(
    bind_rows(control_labels, comparison_labels),
    by = c("Week", "IGF_type")
  )

plot_rescue <- ggplot(
  rescue_plot_data,
  aes(
    x = Week,
    y = Perc.WeightLoss.trend,
    color = IGF_type
  )
) +
  annotate(
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = -Inf,
    ymax = 0,
    fill = "#E87722",
    alpha = 0.12
  ) +
  annotate(
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = 0,
    ymax = Inf,
    fill = "#0C2340",
    alpha = 0.12
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  ) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.15,
    position = position_dodge(width = 0.25)
  ) +
  geom_point(
    size = 4,
    position = position_dodge(width = 0.25)
  ) +
  geom_text(
    aes(
      y = upper.CL + 0.05,
      label = sig_vs_control
    ),
    position = position_dodge(width = 0.25),
    size = 8,
    show.legend = FALSE
  ) +
  annotate(
    "label",
    x = 4.5,
    y = 0.72,
    label = "Energetic Rescue",
    hjust = 0,
    color = "#0C2340",
    fill = "white",
    fontface = "bold",
    size = 6,
    label.size = 0.5,
    label.padding = grid::unit(0.15, "lines")
  ) +
  annotate(
    "label",
    x = 4.5,
    y = -0.45,
    label = "Energetic Trade-off",
    hjust = 0,
    color = "#E87722",
    fill = "white",
    fontface = "bold",
    size = 6,
    label.size = 0.5,
    label.padding = grid::unit(0.15, "lines")
  ) +
  scale_color_manual(
    values = c(
      "IGF1" = "#E87722",
      "IGF2" = "#0C2340",
      "No_IGF" = "grey40"
    ),
    labels = c(
      "IGF1" = "IGF1",
      "IGF2" = "IGF2",
      "No_IGF" = "None"
    )
  ) +
  labs(
    title = "IGF2 Reduces the Constraint on Regeneration",
    subtitle = "Change in regeneration per 1% increase in weight loss",
    x = "Week",
    y = "Slope of Regeneration vs. Weight Loss",
    color = "Supplementation"
  ) +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )

plot_rescue

ggsave(
  filename = "IGF2_energetic_rescue_figure.png",
  plot = plot_rescue,
  width = 7,
  height = 5,
  dpi = 600
)
```

## 6.6 Residual-variance checks among IGF groups

```{r rescue-residual-variance}
model_frame_rescue <- model.frame(model_igf_split)
model_frame_rescue$resid <- residuals(model_igf_split)

leveneTest(resid ~ IGF_type, data = model_frame_rescue)

model_frame_igf1 <- model_frame_rescue %>%
  filter(IGF_type %in% c("No_IGF", "IGF1"))

model_frame_igf2 <- model_frame_rescue %>%
  filter(IGF_type %in% c("No_IGF", "IGF2"))

leveneTest(resid ~ IGF_type, data = model_frame_igf1)
leveneTest(resid ~ IGF_type, data = model_frame_igf2)
```

# 7. Bootstrap Evaluation of the Rescue Effect

The July analysis contains two bootstrap procedures. Both are retained here because they address different purposes:

1. A parametric bootstrap of the full rescue model for visualizing fixed-effect distributions.
2. A late-stage bootstrap used in the sensitivity analysis of weeks 5–8.

## 7.1 Parametric bootstrap of the full rescue model

```{r bootstrap-full-model}
boot_fixed_effects <- function(fit) {
  fixef(fit)
}

bootstrap_full <- bootMer(
  model_igf_split,
  FUN = boot_fixed_effects,
  nsim = 500,
  type = "parametric"
)

bootstrap_full_df <- as.data.frame(bootstrap_full$t)
names(bootstrap_full_df) <- names(fixef(model_igf_split))

bootstrap_full_long <- bootstrap_full_df %>%
  select(
    `Week:Perc.WeightLoss`,
    `Week:Perc.WeightLoss:IGF_typeIGF1`,
    `Week:Perc.WeightLoss:IGF_typeIGF2`
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "term",
    values_to = "estimate"
  )

bootstrap_plot <- ggplot(
  bootstrap_full_long,
  aes(x = estimate, fill = term)
) +
  geom_density(alpha = 0.9) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(
    values = c("grey40", "#E87722", "#0C2340")
  ) +
  labs(
    x = "Bootstrap estimate",
    y = "Density",
    fill = "Model term"
  ) +
  theme_classic() +
  theme(legend.position = "none")

bootstrap_plot

ggsave(
  filename = "bootstrapConfid.png",
  plot = bootstrap_plot,
  width = 2.5,
  height = 2,
  dpi = 600
)
```

# 8. Additional and Sensitivity Analyses

## 8.1 Administered hormone dose

```{r dose-analysis}
igf_data <- data %>%
  filter(Treatment %in% c("DR.IGF1", "DR.IGF2")) %>%
  mutate(
    IGF_type = factor(
      Treatment,
      levels = c("DR.IGF1", "DR.IGF2")
    )
  ) %>%
  group_by(Animal_ID) %>%
  fill(Dose, .direction = "downup") %>%
  ungroup()

model_dose <- lmer(
  Rate_Reg ~ Week * Dose * IGF_type +
    Tail +
    (1 | Animal_ID),
  data = igf_data
)

summary(model_dose)
anova(model_dose)
```

**Interpretation:** Modest individual variation in administered dose did not detectably influence regeneration rate or explain the different responses to IGF1 and IGF2.

## 8.2 Absolute regenerated tail length

```{r absolute-regeneration-rescue}
model_rescue_absolute <- lmer(
  Regeneration ~
    Week * Perc.WeightLoss * IGF +
    Tail +
    (1 | Animal_ID),
  data = data
)

summary(model_rescue_absolute)
anova(model_rescue_absolute)
```

**Manuscript results:**

- Week: \(\beta=3.29\), \(t=22.87\), \(p<0.0001\)
- Week × percent body-mass loss: \(\beta=-0.0166\), \(t=-2.20\), \(p=0.028\)
- Week × percent body-mass loss × IGF: \(\beta=0.0420\), \(t=3.06\), \(p=0.0026\)

## 8.3 Endpoint covariance between energetic state and regeneration

```{r endpoint-tradeoff}
final_regeneration_data <- data %>%
  group_by(Animal_ID) %>%
  summarise(
    final_regeneration = max(Regeneration, na.rm = TRUE),
    weight_loss = first(Perc.WeightLoss),
    Treatment = first(Treatment),
    .groups = "drop"
  )

cor.test(
  final_regeneration_data$weight_loss,
  final_regeneration_data$final_regeneration
)

model_endpoint_tradeoff <- lm(
  final_regeneration ~ weight_loss * Treatment,
  data = final_regeneration_data
)

summary(model_endpoint_tradeoff)
anova(model_endpoint_tradeoff)
```

This endpoint analysis is retained because it demonstrates that the longitudinal mixed-effects model is more sensitive to changes in regenerative dynamics than a single final-time-point comparison.

## 8.4 Final percent body-mass loss among treatment groups

```{r final-weight-loss}
final_weight_loss <- data %>%
  ungroup() %>%
  group_by(Animal_ID) %>%
  summarise(
    final_weight_loss = max(Perc.WeightLoss, na.rm = TRUE),
    Treatment = first(na.omit(Treatment)),
    .groups = "drop"
  ) %>%
  filter(
    !is.na(final_weight_loss),
    is.finite(final_weight_loss),
    !is.na(Treatment)
  ) %>%
  mutate(
    Treatment = droplevels(as.factor(Treatment))
  )

model_final_weight_loss <- lm(
  final_weight_loss ~ Treatment,
  data = final_weight_loss
)

anova(model_final_weight_loss)

final_weight_loss_emmeans <- emmeans(
  model_final_weight_loss,
  pairwise ~ Treatment,
  adjust = "tukey"
)

final_weight_loss_emmeans
```

**Manuscript interpretation:** Final percent body-mass loss did not differ among the three diet-restricted groups, supporting the conclusion that the IGF2 rescue effect was not explained by reduced energetic deficit.

```{r final-weight-loss-plot}
ggplot(
  final_weight_loss,
  aes(
    x = Treatment,
    y = final_weight_loss,
    fill = Treatment
  )
) +
  geom_violin(alpha = 0.3, trim = FALSE) +
  geom_jitter(width = 0.1, alpha = 0.7, size = 2) +
  stat_summary(
    fun = mean,
    geom = "point",
    size = 4,
    color = "black"
  ) +
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    width = 0.15,
    color = "black"
  ) +
  scale_fill_manual(values = treatment_colors) +
  labs(
    x = "Treatment",
    y = "Final percent body-mass loss"
  ) +
  theme_classic() +
  theme(legend.position = "none")
```

# 9. Evaluation of Weeks 5–8

## 9.1 Visual inspection of individual regeneration trajectories

```{r later-timepoint-trajectories}
regeneration_raw <- read.csv(
  "R.analysis.currated.csv",
  stringsAsFactors = FALSE
)

regeneration_after_week_2 <- regeneration_raw %>%
  filter(Week > 2)

plot_length_loss_all <- ggplot(
  regeneration_raw,
  aes(
    x = Week,
    y = Regeneration,
    color = Animal_ID,
    group = Animal_ID
  )
) +
  geom_line() +
  geom_text(
    aes(label = Animal_ID),
    hjust = 0,
    vjust = 0,
    size = 2
  ) +
  facet_wrap(
    ~Treatment,
    scales = "free",
    ncol = 2
  ) +
  theme(legend.position = "none")

plot_length_loss_late <- ggplot(
  regeneration_after_week_2,
  aes(
    x = Week,
    y = Regeneration,
    color = Animal_ID,
    group = Animal_ID
  )
) +
  geom_line() +
  geom_text(
    aes(label = Animal_ID),
    hjust = 0,
    vjust = 0,
    size = 2
  ) +
  facet_wrap(
    ~Treatment,
    scales = "free",
    ncol = 2
  ) +
  theme(legend.position = "none")

plot_length_loss_all
plot_length_loss_late

ggsave(
  filename = "length_loss.png",
  plot = plot_length_loss_all,
  width = 8,
  height = 8,
  dpi = 600
)

ggsave(
  filename = "length_loss_sub.png",
  plot = plot_length_loss_late,
  width = 8,
  height = 8,
  dpi = 600
)
```

## 9.2 Early-only versus full eight-week model

```{r early-versus-full-models}
model_early <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * IGF_type +
    Tail +
    (1 | Animal_ID),
  data = data %>%
    filter(Week <= 4)
)

model_full <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * IGF_type +
    Tail +
    (1 | Animal_ID),
  data = data
)

summary(model_early)
summary(model_full)
```

**Reported comparison:**

- Weeks 1–4: IGF2 interaction \(\beta=0.088\), \(p=0.213\)
- Weeks 1–8: IGF2 interaction \(\beta=0.114\), \(p=0.000138\)

## 9.3 Within-individual repeatability

```{r full-model-icc}
icc(model_full)
```

**Reported result:** Adjusted ICC approximately 0.47, indicating moderate within-individual repeatability across the full eight-week dataset.

## 9.4 Residual variation across weeks

```{r residual-variation-by-week}
full_model_residual_data <- model.frame(model_full) %>%
  mutate(
    residual = residuals(model_full),
    absolute_residual = abs(residual)
  )

model_absolute_residuals <- lm(
  absolute_residual ~ factor(Week),
  data = full_model_residual_data
)

anova(model_absolute_residuals)
```

## 9.5 Late-stage bootstrap

```{r late-stage-bootstrap}
late_data <- data %>%
  filter(Week >= 5)

model_late <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * IGF_type +
    Tail +
    (1 | Animal_ID),
  data = late_data
)

late_boot_function <- function(fit) {
  fixef(fit)["Week:Perc.WeightLoss:IGF_typeIGF2"]
}

bootstrap_late <- bootMer(
  model_late,
  FUN = late_boot_function,
  nsim = 1000
)

late_bootstrap_interval <- quantile(
  bootstrap_late$t,
  c(0.025, 0.5, 0.975),
  na.rm = TRUE
)

late_bootstrap_interval
```

**Reported result:** Median \(=0.142\), 95% bootstrap interval \(=0.051\) to \(0.231\).

# 10. Evaluation of Original Tail Length as a Covariate

Models are fitted with maximum likelihood for valid AIC comparison.

```{r tail-covariate-aic}
model_rescue_without_tail <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * IGF_type +
    (1 | Animal_ID),
  data = data,
  REML = FALSE
)

model_rescue_with_tail <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * IGF_type +
    Tail +
    (1 | Animal_ID),
  data = data,
  REML = FALSE
)

AIC(
  model_rescue_without_tail,
  model_rescue_with_tail
)

summary(model_rescue_with_tail)
anova(model_rescue_with_tail)
```

**Reported result:** Inclusion of original tail length improved model fit by \(\Delta AIC=6.9\), so Tail was retained as an additive covariate in the primary rescue model.

# 11. Publication Figures

## 11.1 Figure 1: dietary restriction and the energetic trade-off

```{r publication-figure-1}
figure1_weight_data <- diet_final_acclimation %>%
  mutate(
    Diet_display = case_when(
      Treatment == "AdLib" ~ "High",
      Treatment %in% c("DR", "DR.IGF1", "DR.IGF2") ~ "Low",
      TRUE ~ NA_character_
    ),
    Diet_display = factor(
      Diet_display,
      levels = c("High", "Low")
    )
  )

figure1_regeneration_data <- data %>%
  mutate(
    Diet_display = case_when(
      Treatment == "AdLib" ~ "High",
      Treatment %in% c("DR", "DR.IGF1", "DR.IGF2") ~ "Low",
      TRUE ~ NA_character_
    ),
    Diet_display = factor(
      Diet_display,
      levels = c("High", "Low")
    )
  )

panel_1a <- ggplot(
  figure1_weight_data,
  aes(x = Diet_display, y = Perc_Orig)
) +
  geom_violin(
    fill = "grey75",
    color = "black",
    trim = FALSE,
    scale = "width",
    alpha = 0.8
  ) +
  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    fill = "white",
    color = "black"
  ) +
  geom_jitter(
    width = 0.08,
    size = 1.8,
    shape = 21,
    fill = "white",
    color = "black",
    alpha = 0.8
  ) +
  labs(
    title = "Dietary restriction increased weight loss",
    x = "Diet",
    y = "Percent of original body mass"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

panel_1b <- ggplot(
  figure1_regeneration_data,
  aes(
    x = Week,
    y = Perc.Regeneration,
    shape = Diet_display,
    linetype = Diet_display,
    group = Diet_display
  )
) +
  stat_summary(
    fun = mean,
    geom = "line",
    linewidth = 1.2,
    color = "black"
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    size = 3,
    color = "black"
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.15,
    color = "black"
  ) +
  scale_shape_manual(
    values = c("High" = 16, "Low" = 17)
  ) +
  scale_linetype_manual(
    values = c("High" = "solid", "Low" = "dashed")
  ) +
  labs(
    title = "Dietary restriction reduced regeneration",
    x = "Week",
    y = "Percent regeneration",
    shape = "Diet",
    linetype = "Diet"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

tradeoff_no_igf <- data %>%
  filter(IGF_type == "No_IGF")

model_tradeoff_no_igf <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss +
    (1 | Animal_ID),
  data = tradeoff_no_igf
)

tradeoff_slopes <- emtrends(
  model_tradeoff_no_igf,
  ~ Week,
  var = "Perc.WeightLoss",
  at = list(Week = c(2, 4, 6, 8))
)

tradeoff_slopes_df <- as.data.frame(
  summary(tradeoff_slopes, infer = TRUE)
) %>%
  mutate(
    sig_zero = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.10  ~ "†",
      TRUE            ~ ""
    )
  )

panel_1c <- ggplot(
  tradeoff_slopes_df,
  aes(
    x = Week,
    y = Perc.WeightLoss.trend
  )
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black",
    linewidth = 0.7
  ) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.18,
    linewidth = 0.8,
    color = "grey30"
  ) +
  geom_point(
    size = 4.5,
    color = "grey30"
  ) +
  geom_text(
    aes(
      y = lower.CL - 0.035,
      label = sig_zero
    ),
    size = 5,
    fontface = "bold",
    color = "grey30"
  ) +
  scale_x_continuous(
    breaks = c(2, 4, 6, 8)
  ) +
  scale_y_continuous(
    breaks = seq(-0.4, 0.1, by = 0.1),
    limits = c(-0.43, 0.15),
    labels = scales::number_format(accuracy = 0.1),
    expand = expansion(mult = c(0.04, 0.05))
  ) +
  labs(
    title = "Energetic trade-off strengthened over time",
    subtitle = "Change in regeneration per 1% increase in weight loss",
    x = "Week",
    y = "Slope of Regeneration\nvs. Weight Loss"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(face = "bold", size = 12),
    axis.title = element_text(face = "bold", size = 13),
    axis.text = element_text(face = "bold", size = 12),
    axis.title.y = element_text(margin = margin(r = 10))
  )

figure_1 <- (panel_1a + panel_1b) / panel_1c +
  plot_layout(heights = c(1, 0.9)) +
  plot_annotation(
    title = "Dietary Restriction Establishes an Energetic Trade-Off",
    tag_levels = "A"
  ) &
  theme(
    plot.tag = element_text(face = "bold", size = 20)
  )

figure_1

ggsave(
  filename = "dietary_restriction_tradeoff_multiplot.png",
  plot = figure_1,
  width = 10,
  height = 7,
  dpi = 600
)
```

## 11.2 Figure 2B: cumulative change in slope

```{r publication-figure-2b}
cumulative_slope_change <- week_specific_slopes_df %>%
  filter(Week %in% c(2, 8)) %>%
  select(
    IGF_type,
    Week,
    Perc.WeightLoss.trend
  ) %>%
  pivot_wider(
    names_from = Week,
    values_from = Perc.WeightLoss.trend,
    names_prefix = "Week_"
  ) %>%
  mutate(
    change = Week_8 - Week_2
  )

plot_cumulative_effect <- ggplot(
  cumulative_slope_change,
  aes(
    x = IGF_type,
    y = change,
    fill = IGF_type
  )
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_col(width = 0.65) +
  scale_fill_manual(values = igf_colors) +
  scale_x_discrete(
    labels = c(
      "No_IGF" = "None",
      "IGF1" = "IGF1",
      "IGF2" = "IGF2"
    )
  ) +
  labs(
    title = "Cumulative Effects",
    x = "Supplementation",
    y = expression(Delta ~ slope ~ relative ~ to ~ control),
    fill = "IGF supplementation"
  ) +
  theme_classic() +
  theme(legend.position = "none")

plot_cumulative_effect

ggsave(
  filename = "IGF2_means.png",
  plot = plot_cumulative_effect,
  width = 2.5,
  height = 2,
  dpi = 600
)
```

## 11.3 Regeneration and body-mass-loss trajectories

```{r regeneration-weightloss-figure}
plot_regeneration_trajectories <- ggplot(
  data,
  aes(
    x = Week,
    y = Perc.Regeneration,
    color = Treatment,
    group = Treatment
  )
) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(fun = mean, geom = "point", size = 2.8) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.15) +
  scale_color_manual(
    values = treatment_colors,
    labels = c(
      "AdLib" = "AdLib",
      "DR" = "DR",
      "DR.IGF1" = "DR + IGF1",
      "DR.IGF2" = "DR + IGF2"
    )
  ) +
  labs(
    title = "Regeneration diverged among treatments",
    x = "Week",
    y = "Percent regeneration",
    color = "Treatment"
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

plot_weightloss_trajectories <- ggplot(
  data,
  aes(
    x = Week,
    y = Perc.WeightLoss,
    color = Treatment,
    group = Treatment
  )
) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(fun = mean, geom = "point", size = 2.8) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.15) +
  scale_color_manual(
    values = treatment_colors,
    labels = c(
      "AdLib" = "AdLib",
      "DR" = "DR",
      "DR.IGF1" = "DR + IGF1",
      "DR.IGF2" = "DR + IGF2"
    )
  ) +
  labs(
    title = "Weight loss did not separate similarly",
    x = "Week",
    y = "Percent weight loss",
    color = "Treatment"
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

combined_regeneration_weightloss <- (
  plot_regeneration_trajectories /
    plot_weightloss_trajectories
) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "IGF2 altered regeneration without broadly reducing energetic deficit"
  ) &
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold")
  )

combined_regeneration_weightloss
```

## 11.4 Restricted-group biology figure

```{r restricted-groups-figure}
restricted_data <- data %>%
  filter(
    Treatment %in% c("DR", "DR.IGF1", "DR.IGF2")
  ) %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("DR", "DR.IGF1", "DR.IGF2")
    )
  )

model_restricted_rescue <- lmer(
  Perc.Regeneration ~
    Week * Perc.WeightLoss * Treatment +
    (1 | Animal_ID),
  data = restricted_data
)

restricted_slopes <- emtrends(
  model_restricted_rescue,
  ~ Treatment | Week,
  var = "Perc.WeightLoss",
  at = list(Week = c(2, 4, 6, 8))
)

restricted_slopes_df <- as.data.frame(
  summary(restricted_slopes, infer = TRUE)
)

restricted_colors <- c(
  "DR" = "grey40",
  "DR.IGF1" = "#E87722",
  "DR.IGF2" = "#0C2340"
)

restricted_labels <- c(
  "DR" = "DR",
  "DR.IGF1" = "DR + IGF1",
  "DR.IGF2" = "DR + IGF2"
)

restricted_panel_a <- ggplot(
  restricted_slopes_df,
  aes(
    x = Week,
    y = Perc.WeightLoss.trend,
    color = Treatment,
    group = Treatment
  )
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.15,
    position = position_dodge(width = 0.25)
  ) +
  geom_point(
    size = 3.5,
    position = position_dodge(width = 0.25)
  ) +
  geom_line(
    linewidth = 1,
    position = position_dodge(width = 0.25)
  ) +
  scale_color_manual(
    values = restricted_colors,
    labels = restricted_labels,
    name = "Treatment"
  ) +
  labs(
    title = "IGF2 changed the regeneration response to weight loss",
    x = "Week",
    y = "Change in regeneration per 1% weight loss"
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

restricted_final_weight_loss <- final_weight_loss %>%
  filter(
    Treatment %in% c("DR", "DR.IGF1", "DR.IGF2")
  ) %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("DR", "DR.IGF1", "DR.IGF2")
    )
  )

restricted_panel_b <- ggplot(
  restricted_final_weight_loss,
  aes(
    x = Treatment,
    y = final_weight_loss,
    fill = Treatment
  )
) +
  geom_violin(trim = FALSE, alpha = 0.25, color = NA) +
  geom_jitter(width = 0.08, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = mean,
    geom = "point",
    size = 3.5,
    color = "black"
  ) +
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    width = 0.15,
    color = "black"
  ) +
  scale_fill_manual(
    values = restricted_colors,
    labels = restricted_labels,
    name = "Treatment"
  ) +
  scale_x_discrete(labels = restricted_labels) +
  labs(
    title = "IGF2 did not reduce body mass loss",
    x = "Treatment",
    y = "Final percent body-mass loss"
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.position = "none"
  )

restricted_biology_figure <- (
  restricted_panel_a |
    restricted_panel_b
) +
  plot_layout(widths = c(1.5, 1)) +
  plot_annotation(
    title = "IGF2 altered regeneration biology without reducing energetic deficit",
    tag_levels = "A"
  ) &
  theme(
    plot.title = element_text(face = "bold"),
    plot.tag = element_text(face = "bold", size = 16)
  )

restricted_biology_figure
```

# 12. Session Information

Including session information documents the versions of R and all attached packages used to render the analysis.

```{r session-information}
sessionInfo()
```
